---
title: "Reproducible Research:-  Project 1 Assignment"
author: "Neil Chapman"
date: "14/01/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load up Libraries

```{r echo=TRUE}
library(dplyr)
library(ggplot2)
```

# Step 1 :- Read in data from activity file, and get summary information on file etc

```{r echo=TRUE}
ds <- read.csv("activity.csv", stringsAsFactors = FALSE)
# get number of rows and columns
dim(ds)
# get first 5 rows of dataset
head(ds, n=5)
# get the last 5 rows of dataset
tail(ds, n=5)
# get info on columns
str(ds)
# get summary information on dataset
summary(ds)
```

# Remove the NA values from dataset

```{r echo=TRUE}
ds_na <- ds[complete.cases((ds)),]
head(ds_na,n=20)
dim(ds_na)
str(ds_na)
```

# Q1 : What is mean total number of steps taken per day?
- Calculate the total number of steps taken per day
- Make a histogram of the total number of steps taken each day
- Calculate and report the mean and median of the total number of steps taken per day

```{r echo=TRUE}
grp <- group_by(ds_na,date)
sum_steps <- summarise(grp,sum(steps))
head(sum_steps,n=5)
tail(sum_steps,n=5)
```

- Make a histogram of the total number of steps taken each day
```{r echo=TRUE}
hist(x=sum_steps$`sum(steps)`,
     col="green",
     breaks=20,
     xlab="Daily - Total Steps",
     ylab="Frequency",
     main="Histogram of Total Number of steps daily")
```

- Calculate and report the mean and median of the total number of steps taken per day
```{r}
# Calculate median of total number of steps per day
summarise(sum_steps,median(sum_steps$`sum(steps)`))
```
Median = 10765
```{r echo=TRUE}
# Calculate mean total number of steps taken per day
summarise(sum_steps,mean(sum_steps$`sum(steps)`))
```
Mean = 10766

# Q2:- What is the average daily activity pattern?

- Make a time series plot (i.e. ðšðš¢ðš™ðšŽ = "ðš•") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
- Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo=TRUE}
grp <- group_by(ds_na,interval)
meansteps <- summarise(grp,mean(steps))
# meansteps$date <- as.Date(meansteps$date,'%Y-%m-%d')

plot(meansteps$interval, 
     meansteps$`mean(steps)`, 
     type = "l", 
     las = 1, 
     col = "red", 
     main = "Avg Steps taken - Time Series",
     col.main = "blue",
     font.main = 4,
     xlab = "Daily Intervals",
     ylab = "Step Averages",
     xlim = c(min(meansteps$interval),max(meansteps$interval)))
```

- Now find out where the maximum value occurred ...
```{r echo=TRUE}
# find out where the maximum value occurred
meansteps[which.max(meansteps$`mean(steps)`),]
```
Maximum average was 206, at interval 835

# Q3 Imputing Missing Values

- Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with ð™½ð™°s)
- Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
- Create a new dataset that is equal to the original dataset but with the missing data filled in.
- Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r echo=TRUE}
nrow(ds)
nrow(ds_na)
```
Therefore the number of missing observations is :- 17568 - 15264 =  2304.

Now addressing parts 2 and 3 of question (devise strategy to fill in missing values, create dataset with missing values filled in)

```{r echo=TRUE}
# create dataframe with all NA values for steps
na_rows <- ds[is.na(ds),]
# create dataframe with all records that do NOT have NA values
notna_rows <- ds[complete.cases(ds),]
# loop through na_rows and assign mean interval value for that day
for(i in 1:nrow(na_rows)) {
    na_rows[i,1] <- meansteps$`mean(steps)`[meansteps$interval==na_rows[i,3]]
}
head(na_rows, n=5)
# now create new dataframe and combine both previous dataframes, na_rows and notna_rows
new_df <- rbind(na_rows,notna_rows)
dim(new_df)
head(new_df, n=5)
tail(new_df, n=5)
```
Now creating a Histogram of total number of steps each day, for this new dataset

```{r}
grp2 <- group_by(new_df,date)
sum_steps_nf <- summarise(grp2,sum(steps))
head(sum_steps_nf,n=5)
tail(sum_steps_nf,n=5)
# create histrogram
hist(x=sum_steps_nf$`sum(steps)`,
     col="red",
     breaks=20,
     xlab="Daily - Total Steps",
     ylab="Frequency",
     main="Histogram of Total Number of steps daily (newly populated DataFrame")
# Calculate median of total number of steps per day
summarise(sum_steps_nf,median(sum_steps_nf$`sum(steps)`))
# Calculate mean total number of steps taken per day
summarise(sum_steps_nf,mean(sum_steps_nf$`sum(steps)`))
```
Setting the NA values to the mean interval for that day, appears to have made the distribution,
(ie. histogram) more of a normal distribution.
Relating to the median + mean values, imnputing the values seems to have had the effect of making them equal.

```{r}

```


